---
title: "Fast and accurate differential transcript usage by testing equivalence class counts"
author:
- Marek Cmero^1
- Nadia Davidson^1†^
- Alicia Oshlack^1,2†^*
output:
  html_document:
    df_print: paged
---

^1^ Murdoch Chidren's Research Institute, Melbourne, Victoria, Australia
^2^ School of Biosciences, Faculty of Science, University of Melbourne,
Melbourne, Victoria, Australia

^†^ These authors contributed equally in supervision of this work.

\* Corresponding author

E-mail: alicia.oshlack@mcri.edu.au

```{r setup, include = FALSE}
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.path     = "cache/",
                      cache.comments = TRUE,
                      echo           = FALSE,
                      error          = FALSE,
                      fig.path       = "figures/",
                      fig.width      = 18,
                      fig.height     = 12,
                      dev            = c('png', 'pdf'),
                      message        = FALSE,
                      warning        = FALSE)
```


```{r libraries}
# plotting
library(ggplot2)
library(gridExtra)
# library(UpSetR)
library(VennDiagram)
library(RColorBrewer)

# DE/DTU
library(edgeR)
library(DEXSeq)
library(DRIMSeq)
library(tximport)

# utils
library(data.table)
library(dplyr)
library(readr)
```


```{r source}
source('R/load_data.R')
source('R/dtu.R')
source('R/util.R')
source('R/plotting.R')
```


```{r options}
options(stringsAsFactors = FALSE)

# globals
thresholds <- c(0.01, 0.05, 0.1)
cutoff <- 0.05
```


```{r load}
# lookup tables to convert salmon transcript IDs to ensembl reference
dm_lookup <- 'ref/drosophila_transcript_id_lookup.txt.gz'
hs_lookup <- 'ref/hsapiens_transcript_id_lookup.txt.gz'
bm_lookup <- 'ref/mm9_tx_lookup.txt.gz'

# ensembl reference to attach gene IDs to trascript IDs
dm_ref <- 'ref/drosophila_transcript_reference.txt.gz'
hs_ref <- 'ref/hsapiens_transcript_reference.txt.gz'
bm_ref <- 'ref/mm9_transcript_reference.txt.gz'

# load EC data from complete matrices
dm_ec <- load_ec_data('data/drosophila/ec_counts/drosophila_ec_matrix.txt.gz', dm_lookup, dm_ref)
hs_ec <- load_ec_data('data/hsapiens/ec_counts/hsapiens_ec_matrix.txt.gz', hs_lookup, hs_ref)
bm_ec <- load_ec_data('data/bottomly/ec_counts/bottomly_ec_counts.txt.gz', bm_lookup, bm_ref)

# load transcript data from salmon quant.sf files
dm_tx <- load_tx_data('data/drosophila/tx_counts/', dm_lookup, dm_ref)
hs_tx <- load_tx_data('data/hsapiens/tx_counts/', hs_lookup, hs_ref)
bm_tx <- load_tx_data('data/bottomly/tx_counts/', bm_lookup, bm_ref)

# load exon data from dexseq counts
dm_ex <- load_ex_data('data/drosophila/exon_counts/', 'Dm')
hs_ex <- load_ex_data('data/hsapiens/exon_counts/', 'Hs')
bm_ex <- load_ex_data('data/bottomly/exon_counts/', 'SRR')

# add species name
hs_ec$species <- 'hsapiens'; hs_tx$species <- 'hsapiens'; hs_ex$species <- 'hsapiens'
dm_ec$species <- 'drosophila'; dm_tx$species <- 'drosophila'; dm_ex$species <- 'drosophila'
bm_ec$species <- 'bottomly'; bm_tx$species <- 'bottomly'; bm_ex$species <- 'bottomly'

# load bottomly group information
# NOTE: download the SraRunTabler from https://www.ncbi.nlm.nih.gov/Traces/study/ for SRP004777
run_tbl <- read.delim('ref/SraRunTable.txt')
bm_samples <- data.frame(sample=run_tbl$Run,
                    type=sapply(run_tbl$Sample_Name, function(x){strsplit(x, '_')[[1]][1]}))
bm_group1 <- bm_samples$sample[bm_samples$type=='B6']
bm_group2 <- bm_samples$sample[bm_samples$type=='D2']
```

# Figure 2

The number of transcripts, equivalence classes and exons per gene (top), where each feature has at least one associated read. The density of the variance of counts over the mean (bottom) for each feature (calculated per condition).

```{r Figure2}
# get count summaries by feature type
feat_summary <- count_features(hs_ec, hs_tx, hs_ex, 'hsapiens', 'Hs')
feat_summary <- rbind(feat_summary, count_features(dm_ec, dm_tx, dm_ex, 'drosophila', 'Dm'))
feat_summary <- rbind(feat_summary, count_features(bm_ec, bm_tx, bm_ex, 'bottomly', 'SRR'))

datasets <- c('drosophila', 'hsapiens', 'bottomly')
features <- c('Transcripts', 'Equivalence classes', 'Exons')
feat_summary$species <- factor(feat_summary$species, levels=datasets)
feat_summary$feature <- factor(feat_summary$feature, levels=features)

# calculate feature variances across groups
feat_var <- calculate_feature_variance(dm_ec, 'Dm')
feat_var <- rbind(feat_var,
                  calculate_feature_variance(dm_tx, 'Dm', feature='Transcripts'))
feat_var <- rbind(feat_var,
                  calculate_feature_variance(dm_ex, 'Dm', feature='Exons'))
feat_var <- rbind(feat_var,
                  calculate_feature_variance(hs_ec, 'Hs', species='hsapiens'))
feat_var <- rbind(feat_var,
                  calculate_feature_variance(hs_tx, 'Hs', species='hsapiens', feature='Transcripts'))
feat_var <- rbind(feat_var,
                  calculate_feature_variance(hs_ex, 'Hs', species='hsapiens', feature='Exons'))
feat_var <- rbind(feat_var,
                  calculate_feature_variance(bm_ec, 'SRR', species='bottomly',
                                             group1=bm_group1, group2=bm_group2))
feat_var <- rbind(feat_var,
                  calculate_feature_variance(bm_tx, 'SRR', species='bottomly', feature='Transcripts',
                                             group1=bm_group1, group2=bm_group2))
feat_var <- rbind(feat_var,
                  calculate_feature_variance(bm_ex, 'SRR', species='bottomly', feature='Exons',
                                             group1=bm_group1, group2=bm_group2))

feat_var$species <- factor(feat_var$species, levels=c('drosophila', 'hsapiens', 'bottomly'))

# plot results
cols <- c('Transcripts' = '#ff7f00',
          'Exons' = '#e41a1c',
          'Equivalence classes' = '#377eb8')
f2a <- ggplot(feat_summary, aes(feature, V1)) +
            geom_boxplot(alpha = 0.6) +
            theme_bw() +
            theme(legend.position = 'none',
                  text = element_text(size = 18)) +
            scale_y_log10(limits=c(1, 1000)) +
            ylab('number per gene') +
            xlab('') +
            facet_wrap(~species)

f2b <- ggplot(feat_var, aes(log2(variance / mean), colour = data)) +
            geom_density(adjust=3) +
            theme_bw() +
            theme(legend.title = element_blank(),
                  legend.position = 'bottom',
                  text = element_text(size = 18)) +
            scale_colour_manual(values = cols) +
            facet_grid(~species)

grid.arrange(f2a, f2b, nrow=2)
```


```{r run-diffsplice}
group <- rep(c('c1','c2'), each=3)
dm_ec_results <- run_diffsplice(dm_ec, group, 'Dm', feature='ec')
dm_tx_results <- run_diffsplice(dm_tx, group, 'Dm', feature='tx')
dm_ex_results <- run_diffsplice(dm_ex, group, 'Dm', feature='ex')

hs_ec_results <- run_diffsplice(hs_ec, group, 'Hs', feature='ec')
hs_tx_results <- run_diffsplice(hs_tx, group, 'Hs', feature='tx')
hs_ex_results <- run_diffsplice(hs_ex, group, 'Hs', feature='ex')

sample_order <- colnames(bm_ec[,.SD,.SDcols = names(bm_ec) %like% 'SRR'])
group <- as.numeric(sample_order %in% bm_group1)
bm_ec_results <- run_diffsplice(bm_ec, group, 'SRR', feature='ec')
bm_tx_results <- run_diffsplice(bm_tx, group, 'SRR', feature='tx')
bm_ex_results <- run_diffsplice(bm_ex, group, 'SRR', feature='ex')
```

# Supplementary Figure 1

Shows the dispersion versus mean normalised counts for all features across the three data sets, generated using DEXSeq's `plotDispEsts` function. As described in Love et al.[1], the red line shows the fitted dispersion-mean trend, the blue dots indicate the shrunken dispersion estimates, and the blue circles indicate outliers not shrunk towards the prior.

```{r SupplementaryFigure1, fig.width=15, fig.height=15, dev='png', pointsize=24}
par(mfrow=c(3,3))
plotDispEsts(dm_tx_results[['dexseq_object']], xlim=c(1,5e5)); title('Drosophila Transcripts')
plotDispEsts(dm_ec_results[['dexseq_object']], xlim=c(1,5e5)); title('Drosophila Equivalence classes')
plotDispEsts(dm_ex_results[['dexseq_object']], xlim=c(1,5e5)); title('Drosophila Exons')
plotDispEsts(hs_tx_results[['dexseq_object']], xlim=c(1,5e5)); title('Hsapiens Transcripts')
plotDispEsts(hs_ec_results[['dexseq_object']], xlim=c(1,5e5)); title('Hsapiens Equivalence classes')
plotDispEsts(hs_ex_results[['dexseq_object']], xlim=c(1,5e5)); title('Hsapiens Exons')
plotDispEsts(bm_tx_results[['dexseq_object']], xlim=c(1,5e5)); title('Bottomly Transcripts')
plotDispEsts(bm_ec_results[['dexseq_object']], xlim=c(1,5e5)); title('Bottomly Equivalence classes')
plotDispEsts(bm_ex_results[['dexseq_object']], xlim=c(1,5e5)); title('Bottomly Exons')
```

Added: Lancaster aggregation of the equivalence class p-values to get gene p-values, followed by correction to get gene FDR-adjusted p-values.

```{r}
library(aggregation)
library(dplyr)
dm_dexseq_results <- as.data.frame(DEXSeqResults(dm_ec_results$dexseq_object))
dm_lancaster_results <- dm_dexseq_results %>% group_by(groupID) %>% summarize(pval = lancaster(pvalue, log(exonBaseMean)))
dm_lancaster_results$gene_FDR <- p.adjust(dm_lancaster_results$pval, 'BH')
dm_lancaster_results <- data.frame(gene = dm_lancaster_results$groupID,
                                      FDR = dm_lancaster_results$gene_FDR)

hs_dexseq_results <- as.data.frame(DEXSeqResults(hs_ec_results$dexseq_object))
hs_lancaster_results <- hs_dexseq_results %>% group_by(groupID) %>% summarize(pval = lancaster(pvalue, log(exonBaseMean)))
hs_lancaster_results$gene_FDR <- p.adjust(hs_lancaster_results$pval, 'BH')
hs_lancaster_results <- data.frame(gene = hs_lancaster_results$groupID,
                                      FDR = hs_lancaster_results$gene_FDR)
```


```{r Figure3a}
# NOTE: obtain truth data from http://imlspenticton.uzh.ch/robinson_lab/splicing_comparison/

results <- NULL
results[['dexseq_equivalence_class']] <- dm_ec_results[['gene_FDR']]
results[['dexseq_salmon']] <- dm_tx_results[['gene_FDR']]
results[['dexseq_exons']] <- dm_ex_results[['gene_FDR']]

results[['lancaster_equivalence_class']] <- dm_lancaster_results

truth <- read.delim('ref/soneson_results/truth_drosophila_non_null_missing20_ms.txt')
test <- read.delim('ref/soneson_results/merged_results_all_drosophila.txt')
test_reduced <- test[,grep('gene|featurecounts_flat|kallisto', colnames(test))]

res <- get_fdr_tpr_stats(test_reduced, truth, results, thresholds, 'drosophila')

results <- NULL
results[['dexseq_equivalence_class']] <- hs_ec_results[['gene_FDR']]
results[['dexseq_salmon']] <- hs_tx_results[['gene_FDR']]
results[['dexseq_exons']] <- hs_ex_results[['gene_FDR']]

results[['lancaster_equivalence_class']] <- hs_lancaster_results

truth <- read.delim('ref/soneson_results/truth_human_non_null_missing20_ms.txt')
test <- read.delim('ref/soneson_results/merged_results_all_human.txt')
test_reduced <- test[,grep('gene|featurecounts_flat|kallisto', colnames(test))]

res <- rbind(res,
             get_fdr_tpr_stats(test_reduced, truth, results, thresholds, 'hsapiens'))

res$method <- gsub('.adjP', '', res$method)
res$method <- gsub('dexseq_', '', res$method)
res$method[res$method=='exons'] <- 'dexseq_count_exons'

f3a <- ggplot(res, aes(FDR, TPR, group=method, colour=method)) +
        geom_line(size=0.5) +
        geom_point(size=2, shape=1, stroke=1) + theme_bw() + ylim(0,1) + xlim(0,1) +
        geom_vline(xintercept = thresholds,
                   colour='grey',
                   linetype='dotted') + facet_wrap(~species) +
        theme(legend.position = 'bottom',
              legend.title = element_blank(),
              text = element_text(size = 18)) +
        scale_color_brewer(palette = 'Set1')

f3a
```




# Supplementary Figure 2

Shows the significant genes (FDR < 0.05) shared between the methods, obtained from DEXSeq run on the full Bottomly et al.[2] data set for each feature.

```{r SupplementaryFigure2, fig.width=6, fig.height=6}
genes_ec <- distinct(bm_ec[,.SD,.SDcols = names(bm_ec) %in% c('gene_id', 'ec_names')])
genes_tx <- distinct(bm_tx[,.SD,.SDcols = names(bm_tx) %in% c('gene_id', 'ensembl_id')])
genes_ex <- distinct(bm_ex[,.SD,.SDcols = names(bm_ex) %in% c('gene_id', 'exon_id')])
colnames(genes_ec)[1] <- 'feature_id'
colnames(genes_tx)[1] <- 'feature_id'
colnames(genes_ex)[1] <- 'feature_id'

counts_ec <- distinct(bm_ec[,.SD,.SDcols = names(bm_ec) %like% 'SRR|gene_id|ec_names'])
counts_tx <- distinct(bm_tx[,.SD,.SDcols = names(bm_tx) %like% 'SRR|gene_id|ensembl_id'])
counts_ex <- distinct(bm_ex[,.SD,.SDcols = names(bm_ex) %like% 'SRR|gene_id|exon_id'])

bm_qec <- bm_ec_results[['gene_FDR']]
bm_qtx <- bm_tx_results[['gene_FDR']]
bm_qex <- bm_ex_results[['gene_FDR']]

true_ec <- unique(bm_qec[bm_qec$FDR<cutoff,]$gene)
true_tx <- unique(bm_qtx[bm_qtx$FDR<cutoff,]$gene)
true_ex <- unique(bm_qex[bm_qex$FDR<cutoff,]$gene)

dt_truth <- NULL
dt_truth[['Equivalence classes']] <- true_ec
dt_truth[['Transcripts']] <- true_tx
dt_truth[['Exons']] <- true_ex

# upset(fromList(dt_truth), text.scale = 2)

cols <- brewer.pal(3, 'Set3')
venn.plot <- venn.diagram(dt_truth,
                          NULL,
                          fill=cols,
                          alpha=c(0.5,0.5,0.5),
                          cex = 2,
                          cat.fontface=1,
                          category.names=names(dt_truth),
                          main="", scaled=FALSE)
grid.draw(venn.plot)
```

# Figure 3

Shows the performance of the equivalence class method for differential transcript usage, compared to other state-of-the-art methods described in Soneson et al.[3] (top). The bottom plots show the ability of the equivalence class, transcript and exon-based methods to recreate the results of a full comparison (10 vs. 10) of the Bottomly data, using only a (randomly selected) subset of samples (3 vs. 3) across 20 iterations. The union of all genes called as significant across all three methods is used for the FDR, and the intersect (genes called by all three methods) is used for the TPR. Full results (union, intersect and each method's individual truth set) is shown in Supplementary Figure 3.

```{r Figure3, fig.width=12, fig.height=10.5}
# Bottomly data subset testing
N = 3
iters = 20
res_ec <- NULL
res_tx <- NULL
res_ex <- NULL

# seeds used for paper analysis
seeds <- c(7601, 6989, 3551, 2774, 1389,
           1471, 6763, 5167, 3342, 1642,
           6589, 9151, 4694, 1917, 4324,
           5513, 2414, 1424, 6536, 7624)

for(i in 1:iters) {
    set.seed(seeds[i])
    comp <- get_random_comp(bm_samples, N)

    # equivalence classes
    counts <- data.frame(counts_ec[,.SD,.SDcols = names(counts_ec) %in% comp$sample])
    res_ec[[i]] <- run_dexseq(counts, genes_ec, comp$type, 0, 0)[['gene_FDR']]; gc()

    # transcripts
    counts <- data.frame(counts_tx[,.SD,.SDcols = names(counts_tx) %in% comp$sample])
    counts <- sapply(counts, round)
    res_tx[[i]] <- run_dexseq(counts, genes_tx, comp$type, 0, 0)[['gene_FDR']]; gc()

    # exons
    counts <- data.frame(counts_ex[,.SD,.SDcols = names(counts_ex) %in% comp$sample])
    res_ex[[i]] <- run_dexseq(counts, genes_ex, comp$type, 0, 0)[['gene_FDR']]; gc()
}

cols <- c('Transcripts' = '#ff7f00',
          'Equivalence classes' = '#377eb8',
          'Exons' = '#e41a1c')
results <- get_subset_tests_results(res_ec, res_tx, res_ex, true_ec, true_tx, true_ex)
f3b <- plot_bottomly_boxplot(results, cols, title='FDR (union)', toplot='FDR', hline=cutoff)
f3c <- plot_bottomly_boxplot(results, cols, title='TPR (intersect)', toplot='TPR')

grid.arrange(f3a, f3b, f3c, layout_matrix=rbind(c(1,1), c(2,3)), nrow=2)
```

# Supplementary Figure 3

Shows the ability of the three methods to recreate the results of a full comparison (10 vs. 10) of the Bottomly et al.[2] data using random subsets of 3 vs. 3 samples across 20 iterations. The lines between the plots join data points from the same iteration. Each row uses a different 'truth' set: union is the set of genes called significant by any method, intersect is the set of genes called significant by all methods, and individual is the set of genes called significant by that method only.

```{r SupplementaryFigure3, fig.width=12, fig.height=15.75}
results <- get_subset_tests_results(res_ec, res_tx, res_ex,
                                    true_ec, true_tx, true_ex,
                                    method = c(FDR='union', TPR='union'))
s3a <- plot_bottomly_boxplot(results, cols, title='FDR (union)', toplot='FDR', hline=cutoff, lines=T)
s3b <- plot_bottomly_boxplot(results, cols, title='TPR (union)', toplot='TPR', lines=T)

results <- get_subset_tests_results(res_ec, res_tx, res_ex,
                                    true_ec, true_tx, true_ex,
                                    method = c(FDR='intersect', TPR='intersect'))
s3c <- plot_bottomly_boxplot(results, cols, title='FDR (intersect)', toplot='FDR', hline=cutoff, lines=T)
s3d <- plot_bottomly_boxplot(results, cols, title='TPR (intersect)', toplot='TPR', lines=T)

results <- get_subset_tests_results(res_ec, res_tx, res_ex,
                                    true_ec, true_tx, true_ex,
                                    method = c(FDR='individual', TPR='individual'))
s3e <- plot_bottomly_boxplot(results, cols, title='FDR (individual)', toplot='FDR', hline=cutoff, lines=T)
s3f <- plot_bottomly_boxplot(results, cols, title='TPR (individual)', toplot='TPR', lines=T)

grid.arrange(s3a, s3b,
             s3c, s3d,
             s3e, s3f, ncol=2)
```

# Supplementary Figure 4

The number of false positives versus each gene's rank (by FDR) for one iteration (3 vs. 3) of the Bottomly subset tests for the top 500 genes. The union of significant genes across all methods was used as the truth set.

```{r SupplementaryFigure4, fig.width=10.5, fig.height=7.5}
pick_iter <- 1
true_dtu <- union(union(true_ec, true_tx), true_ex)
false_dtu <- union(union(genes_ec$gene_id, genes_tx$gene_id), genes_ex$gene_id)
false_dtu <- false_dtu[!false_dtu%in%true_dtu]

rp <- get_rank_orders(res_ec, false_dtu, 'Equivalence classes')
rp <- rbind(rp, get_rank_orders(res_tx, false_dtu, 'Transcripts'))
rp <- rbind(rp, get_rank_orders(res_ex, false_dtu, 'Exons'))

ggplot(rp, aes(gene_rank, false_positives, colour = feature)) +
    geom_line() +
    theme_bw() +
    xlim(0,500) +
    ylim(0,500) +
    scale_color_manual(values = cols) +
    ylab('False positives') +
    xlab('Gene rank') +
    theme(legend.title = element_blank(),
          text = element_text(size = 18))
```

# Supplementary Figure 5

Kallisto[4] versus salmon's[5] performance on the Bottomly subset testing experiments, using each method's significant genes from the full (10 vs. 10) run as the truth set for calculating both metrics.

```{r SupplementaryFigure5, fig.width=12, fig.height=5.25}
# load in kallisto data
bm_kec <- load_ec_data('data/bottomly/ec_counts/bottomly_kallisto_ec_counts.txt.gz', NA, bm_ref, salmon=F)
counts_kec <- distinct(bm_kec[,.SD,.SDcols = names(bm_kec) %like% 'SRR|gene_id|ec_names'])
genes_kec <- distinct(bm_kec[,.SD,.SDcols = names(bm_kec) %in% c('gene_id', 'ec_names')])
colnames(genes_kec)[1] <- 'feature_id'

# run diffsplice
sample_order <- colnames(bm_kec[,.SD,.SDcols = names(bm_kec) %like% 'SRR'])
group <- as.numeric(sample_order %in% bm_group1)
bm_kec_results <- run_diffsplice(bm_kec, group, 'SRR', feature='ec')

# run kallisto iteration tests on Bottomly data (using same seeds as before)
for(i in 1:iters) {
    set.seed(seeds[i])
    comp <- get_random_comp(bm_samples, N)

    counts <- data.frame(counts_kec[,.SD,.SDcols = names(counts_kec) %in% comp$sample])
    res_kec[[i]] <- run_dexseq(counts, genes_kec, comp$type, 0, 0)[['gene_FDR']]; gc()
}

# calculate kallisto results
kal_bm_qec <- bm_kec_results[['gene_FDR']]
true_kec <- unique(kal_bm_qec[kal_bm_qec$FDR<cutoff,]$gene)
kal_res <- get_subset_tests_results(res_kec, res_tx, res_ex,
                                    true_kec, true_tx, true_ex,
                                    method = c(FDR='individual', TPR='individual'))
kal_res <- kal_res[kal_res$feature %in% 'Equivalence classes', ]

# use last salmon results for comparison (individual FDR/TPR calculation)
sal_res <- results[results$feature %in% 'Equivalence classes', ]

# set up results for plotting
kal_res <- melt(kal_res[,c('iter', 'FDR', 'TPR')], id.vars='iter', value.name='kallisto')
sal_res <- melt(sal_res[,c('iter', 'FDR', 'TPR')], id.vars='iter', value.name='salmon')
res <- merge(sal_res, kal_res, by=c('iter', 'variable'))

ggplot(res, aes(salmon, kallisto)) +
    geom_abline(slope = 1, intercept = 0, colour = 'grey') +
    geom_point() +
    theme_bw() +
    xlim(0,1) + ylim(0,1) +
    facet_wrap(~variable)
```

# References

[1] Love, M. I., Huber, W., & Anders, S. (2014). Moderated estimation of fold change and dispersion for RNA-seq data with DESeq2. Genome Biology, 15(12), 1–21. https://doi.org/10.1186/s13059-014-0550-8

[2] Bottomly, D., Walter, N. A. R., Hunter, J. E., Darakjian, P., Kawane, S., Buck, K. J., … Hitzemann, R. (2011). Evaluating gene expression in C57BL/6J and DBA/2J mouse striatum using RNA-Seq and microarrays. PLoS ONE, 6(3). https://doi.org/10.1371/journal.pone.0017820

[3] Soneson, C., Matthes, K. L., Nowicka, M., Law, C. W., & Robinson, M. D. (2016). Isoform prefiltering improves performance of count-based methods for analysis of differential transcript usage. Genome Biology, 17(1), 1–15. https://doi.org/10.1186/s13059-015-0862-3

[4] Bray, N. L., Pimentel, H., Melsted, P., & Pachter, L. (2016). Near-optimal probabilistic RNA-seq quantification. Nature Biotechnology, 34(5), 525–527. https://doi.org/10.1038/nbt.3519

[5] Patro, R., Duggal, G., Love, M. I., Irizarry, R. A., & Kingsford, C. (2017). Salmon provides fast and bias-aware quantification of transcript expression. Nature Methods, 14(4), 021592. https://doi.org/10.1038/nmeth.4197
